---
title: "IV Application (II)"
subtitle: "College proximity as IVs for education"
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(knitr)
library(magrittr)
library(tibble)
```


## Case Description

Research Interests:

With data set `Card1995.dta`, researchers were interest in the return (`log(Wage)`) to education (`edu`) .

In the wage example (`wage`, or log wage `lwage`), Let's consider some of the variables shown below (see @tbl-vars ).

```{r}
#| label: tbl-vars
#| echo: false

vars_sel <- c(
  "obs","lwage",
  "edu", 
  "exp", "exp2",
  "black", "south",  "urban",
  "college",  
  "public", "private",
  "age", "age2",
  "momedu", "dadedu")

label_def <-c(
  "index",
  "quantity variable: log of wage",
  "quantity variable: education years", 
  "quantity variable: working years", "quantity variable: square working years/100",
  "dummy: 1=black; 0=nonblack", 
  "dummy: 1=southern area; 0= other",  
  "dummy: 1=live in urban; 0= other",
  "dummy: 1=college nearby; 0= other",  
  "dummy: 1=public college nearby; 0= other",
  "dummy: 1=private college nearby; 0= other",
  "quantity variable: age (years)", "quantity variable: age square /100",
  "quantity variable: mother' education years", 
  "quantity variable: father' education years")

label_eng <-c(vars_sel)
df_vars <- tibble("variable" = label_eng,
                  "definition" = label_def)

df_vars %>%
  knitr::kable(., align = 'c', caption = 'variables and definition')
```

## Reproducible Sources

**Hansen, B.  Econometrics[M].** Princeton: Princeton University Press, 2022. **Chapter 12:** Instrumental Variables. 

- Table 12.1: Instrumental Variable Wage Regressions

- Table 12.2: Reduced Form Regressions

## Learning Targets

a. Understand the nature of `Endogeneity`.

b. Know the steps of running TSLS method.

c. Be familiar with R package function `systemfit::systemfit()` and `ARE::ivreg()`.

## Exercise Materials 

You can find all the exercise materials in this project under the file directory:

```{r}
#| echo: false
# show files tree
fs::dir_tree(path = here::here("IV-wage-card"),
             recurse = TRUE)
```