---
title: "IV Application (card)"
subtitle: "College proximity as IVs for education"
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(knitr)
library(magrittr)
library(tibble)
```

## Case Description

Research Interests:

With data set `Card1995.dta`, researchers were interest in the return (`log(Wage)`) to education (`edu`) .

In the wage example (`wage`, or log wage `lwage`), Let's consider some of the variables shown below (see @tbl-vars ).

```{r}
#| label: tbl-vars
#| echo: false

vars_sel <- c(
  "obs","lwage",
  "edu", 
  "exp", "exp2",
  "black", "south",  "urban",
  "college",  
  "public", "private",
  "age", "age2",
  "momedu", "dadedu")

label_def <-c(
  "index",
  "quantity variable: log of wage",
  "quantity variable: education years", 
  "quantity variable: working years", "quantity variable: square working years/100",
  "dummy: 1=black; 0=nonblack", 
  "dummy: 1=southern area; 0= other",  
  "dummy: 1=live in urban; 0= other",
  "dummy: 1=college nearby; 0= other",  
  "dummy: 1=public college nearby; 0= other",
  "dummy: 1=private college nearby; 0= other",
  "quantity variable: age (years)", "quantity variable: age square /100",
  "quantity variable: mother' education years", 
  "quantity variable: father' education years")

label_eng <-c(vars_sel)
df_vars <- tibble("variable" = label_eng,
                  "definition" = label_def)

df_vars %>%
  knitr::kable(., align = 'c', caption = 'variables and definition')
```

## Models

### Origin model

The origin model is

$$
\begin{aligned}
  lwage & = \hat{\alpha}_0 +\hat{\alpha}_1 {educ} + \hat{\alpha}_3 exp +\hat{\alpha}_4 exp2 \\
  & +\hat{\alpha}_5 black +\hat{\alpha}_6 south +\hat{\alpha}_7 urban + u_i  
\end{aligned}
$$

### TSLS1: `edu` V.S. `college`

we will use `college` as instruments for `educ` in our IV model setting.

$$\begin{cases}
  \begin{align}
  {edu} &= \hat{\gamma}_0 +\hat{\gamma}_1exp + \hat{\gamma}_2exp2 + \hat{\gamma}_3black + \hat{\gamma}_4south + \hat{\gamma}_5urban + \hat{\theta}_1college +v_i && \text{(stage 1)}\\
  lwage & = \hat{\eta}_1 +\hat{\eta}_2\widehat{edu} + \hat{\eta}_3exp +\hat{\eta}_4exp2 +\hat{\eta}_5 black +\hat{\eta}_6 south +\hat{\eta}_7 urban+ e_i  && \text{(stage 2)}
  \end{align}
\end{cases}$$

### TSLS2: (`edu`,`exp`,`exp2`)  V.S. (`college`, `age`, `age2`)

we use (`college`, `age`, `age2`) as instruments for (`edu`,`exp`,`exp2`) in our IV model setting.

$$\begin{cases}
  \begin{align}
  {edu} &= \hat{\gamma}_0 +\hat{\gamma}_1age + \hat{\gamma}_2age2 + \hat{\gamma}_3black + \hat{\gamma}_4south + \hat{\gamma}_5urban + \hat{\theta}_1college +v_{1i} && \text{(1 of stage 1)}\\
  {exp} &= \hat{\lambda}_0 +\hat{\lambda}_1age + \hat{\lambda}_2age2 + \hat{\lambda}_3black + \hat{\lambda}_4south + \hat{\lambda}_5urban + \hat{\lambda}_1college +v_{2i} && \text{(2 of stage 1)}\\
  {exp2} &= \hat{\delta}_0 +\hat{\delta}_1age + \hat{\delta}_2age2 + \hat{\delta}_3black + \hat{\delta}_4south + \hat{\delta}_5urban + \hat{\delta}_1college +v_{3i} && \text{(3 of stage 1)}\\
  lwage & = \hat{\eta}_1 +\hat{\eta}_2\widehat{edu} + \hat{\eta}_3exp +\hat{\eta}_4exp2 +\hat{\eta}_5 black +\hat{\eta}_6 south +\hat{\eta}_7 urban+ e_i  && \text{(stage 2)}
  \end{align}
\end{cases}$$

### TSLS3: `edu` V.S. (`public`,`private`)

we use both (`public`,`private`) as instruments for `educ` in our IV model setting.

$$\begin{cases}
  \begin{align}
  {edu} &= \hat{\gamma}_0 +\hat{\gamma}_1exp + \hat{\gamma}_2exp2 + \hat{\gamma}_3black + \hat{\gamma}_4south + \hat{\gamma}_5urban + \hat{\theta}_1public + \hat{\theta}_2private +v_i && \text{(stage 1)}\\
  lwage & = \hat{\eta}_1 +\hat{\eta}_2\widehat{edu} + \hat{\eta}_3exp +\hat{\eta}_4exp2 +\hat{\eta}_5 black +\hat{\eta}_6 south +\hat{\eta}_7 urban+ e_i  && \text{(stage 2)}
  \end{align}
\end{cases}$$

### TSLS4: (`edu`, `exp`,`exp2`) V.S. (`public`,`private`,`age`,`age2`)

we use both (`public`,`private`,`age`,`age2`) as instruments for (`edu`, `exp`,`exp2`) in our IV model setting.

$$\begin{cases}
  \begin{align}
    {edu} &= \hat{\gamma}_0 +\hat{\gamma}_1age + \hat{\gamma}_2age2 + \hat{\gamma}_3black + \hat{\gamma}_4south + \hat{\gamma}_5urban + \hat{\theta}_1public + \hat{\theta}_2private +v_{1i} && \text{(1 of stage 1)}\\
  {exp} &= \hat{\lambda}_0 +\hat{\lambda}_1age + \hat{\lambda}_2age2 + \hat{\lambda}_3black + \hat{\lambda}_4south + \hat{\lambda}_5urban + \hat{\lambda}_1public + \hat{\lambda}_2private +v_{2i} && \text{(2 of stage 1)}\\
  {exp2} &= \hat{\delta}_0 +\hat{\delta}_1age + \hat{\delta}_2age2 + \hat{\delta}_3black + \hat{\delta}_4south + \hat{\delta}_5urban + \hat{\delta}_1public + \hat{\delta}_2private +v_{3i} && \text{(3 of stage 1)}\\
  lwage & = \hat{\eta}_1 +\hat{\eta}_2\widehat{edu} + \hat{\eta}_3exp +\hat{\eta}_4exp2 +\hat{\eta}_5 black +\hat{\eta}_6 south +\hat{\eta}_7 urban+ e_i  && \text{(stage 2)}
  \end{align}
\end{cases}$$

## Reproducible Sources

**Hansen, B.  Econometrics[M].** Princeton: Princeton University Press, 2022. **Chapter 12:** Instrumental Variables. 

- Table 12.1: Instrumental Variable Wage Regressions

- Table 12.2: Reduced Form Regressions


## Learning Targets

a. Understand the nature of `Endogeneity`.

b. Know the steps of running TSLS method.

c. Be familiar with R package function `systemfit::systemfit()` and `ARE::ivreg()`.

## Exercise Materials 

You can find all the exercise materials in this project under the file directory:

```{r}
#| echo: false
# show files tree
fs::dir_tree(path = here::here("IV-wage-card"),
             recurse = 1)
```