---
title: "Requirements"
---

## Step 1/3：Prepare Software 

You need install two software (Both can download from <https://posit.co/download/rstudio-desktop/>):

- `R` is an open source (FREE) programming language.

- `Rstudio` (Also FREE) is an IDE (Integrated Development Environment) which will help you coding more efficiently.  



## Step 2/3：Following Course Project

I create a course project of "Econometrics III Accompany", with which students in my " Advanced Econometrics" class can do some programming exercises and make class examples reproducible.

::: {.callout-note}

- This project is under building and will iterate regularly.

- I create a project repository(<https://github.com/huhuaping/course-emiii-accompany>) and host it at <https://github.com/> publicly.

- For those students in china, you can also access the synchronous repository (<https://gitee.com/kevinhhp/course-emiii-accompany>).

:::

There are two ways to get these project files into you PC/Laptop:

- (Recommended) You can download the **".zip"** file by click "Download" button after you click the URL show before.

- Use tools like `gitbash` to **clone** or **fork** the whole project.


Finally, the project directory structure will be shows as bellow:

```{r}
#| echo: false
# show files tree
#fs::dir_ls(path = "D:/github/course-emiii-accompany")
fs::dir_tree(path = "D:/github/course-emiii-accompany",
             recurse = 1)
```


## Step 3/3：Install R packages

R is a function-based object-oriented programming language. When executing R code commands, various R packages are often used to achieve specific programming goals. 

After you have downloaded the project file to your local computer (Step 2/3), you can open the project with Rstudio (Step 1/3), then open 'R/load-pkg-basic.R'(see below) to see the relevant R packages needed for this project and install them.

```{r filename="R/load-pkg-basic.R"}
#| eval: false
#| echo: true

# load necessary R  packages

# system
library(rmarkdown)
library(renv)       # package management
library(yaml)       # quarto project
library(fs)         # file system management
library(here)       # relative path

# basic pkgs
library(tidyverse)  # most useful
library(magrittr)   # pipe %>%
library(knitr)      # R render engine

# read file
library(openxlsx)  # for excel file
library(foreign)   # for stata file

# viz
library(ggplot2)       # graph
library(DT)            # DT table
library(scales)        # tidy numbers
library(webshot2)      # snapshot
library(htmlwidgets)   # interactivity
#library(fontawesome)

# github repo
#renv::install("KWB-R/kwb.utils")
#renv::install("huhuaping/xmerit")  # developed by `Huhuaping`
#library(xmerit)                    # export Latex equation
```

::: {.callout-tip}

To put it simply, R packages come from a few main sources:

- The base package by the R core team

- R packages developed by the R community (widely published on github and other hosting platforms)

- Other source of R packages

:::

How to install these R packages?

Solution 1 (**Not recommend**): use `install.packages("foo")` to install all packages one by one. 

Solution 2 (**Strongly recommend**) :  use `renv` package management tool chain!

- Open the course project with `Rstudio` (double click `course-emiii-accompany.Rproj`). Then you can write and run R command in panel `Console` below the software window.

- First, write and run R command `install.packages("renv")` 

- Then, write and run R command `library(renv)`

- Finally, write and run R command `renv::active()` and `renv::restore()`
